---
title: "bayes_notebook"
author: "Tara Skiba"
date: "2025-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0.1 Load packages

```{r}
library(cmdstanr)
library(rethinking)
library(ggplot2)
library(dplyr)
library(readxl)
library(data.table)
library(patchwork)
library(tidyr)
library(foreach)
library(doParallel)
set.seed(123)
```
Make clusters
```{r}
no_cores <- detectCores() - 3
cl <- makeCluster(no_cores, type = "PSOCK")
registerDoParallel(cl)
```

## 0.2 Load data
```{r}
final_plot_data <- read.csv("tree_plot_data_final_carbon.csv")
final_plot_data <- final_plot_data %>% 
  filter(ownership != "New River Gorge")
ownership_index <- c("Chattahoochee-Oconee", "Cherokee", "Nantahala", "Pisgah","George Washington-Jefferson", "Daniel Boone","Monongahela","Wayne","Great Smoky Mountains","Shenandoah")
 
```

```{r}
selected_variables <- c("cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")
selected_corr_data <- final_plot_data[,selected_variables]

hist_data <- selected_corr_data %>% gather() %>% head()
ggplot(gather(selected_corr_data), aes(value)) + 
  geom_histogram(bins = 10) + 
  facet_wrap(~key, scales = 'free_x')
#final_plot_data[,c("sdi", "tph")] <- scale(final_plot_data[,c("sdi", "tph")])

obs_count <- final_plot_data %>% 
  group_by(ownership) %>% 
  count()
```

## 0.3 Spare functions to call
```{r}

bayesian_rsq = function(mod, obs){
	# Compute the Bayesian R2 for a model

	pred = link(mod, n=10000)
	all_resid = t(obs - t(pred))
	resid_var = apply(all_resid, 1, var)
	mean_var =  apply(pred, 1, var)
	R2 = mean_var / (mean_var + resid_var)
	return(R2)

}

bayesian_rsq_out = function(mod, obs, new_values){
  pred = link(mod, data=new_values,  n=10000)
  all_resid = t(obs - t(pred))
  resid_var = apply(all_resid, 1, var)
  mean_var =  apply(pred, 1, var)
  R2 = mean_var / (mean_var + resid_var)
  return(R2)
}

lppd_newdata = function(mod, newdata, n=1e4){
	# Function to compute lppd on new data

	ll <- sim(mod, data=newdata, ll = TRUE)
	n <- ncol(ll)
	ns <- nrow(ll)
	f <- function(i) log_sum_exp(ll[, i]) - log(ns)
	lppd <- sapply(1:n, f)
	return(lppd)

}

```

## 1.1 LOGO-CV applied on plot level
```{r}
i = 9

final_plot_data_subset <- final_plot_data %>% 
  filter(ownership == ownership_index[i]) %>% 
  filter(fortypcd < 600 & fortypcd >= 500)

plot_IDs <- unique(final_plot_data_subset$uniq_spl_unit_id)

subset = final_plot_data_subset
subset$sqrtcpa = sqrt(final_plot_data_subset$cpa)

#subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] = scale(subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")])

cross_v <- foreach(p = 1:length(plot_IDs), .packages = c("rethinking"))%dopar%{ #    .errorhandling = 'remove'
  id = plot_IDs[p]
  
  boot_data = subset[which(subset$uniq_spl_unit_id != id),]
  test_data = subset[which(subset$uniq_spl_unit_id == id),]
  
  boot_data = boot_data[, c( "sqrtcpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] #"fortypcd",
  test_data = test_data[, c( "sqrtcpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")]

  # Specify model
  mod1 = alist(
    sqrtcpa ~ dnorm(mu, sigma),
    mu <- b0  + b_qmd*qmd + b_baph*baph + b_tph*tph + b_dia*mean_dia + #+ alpha[fortypcd]
      b_ht*mean_actualht + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    b0 ~ dnorm(8, 4),
    sigma ~ dexp(1),
    c(b_qmd, b_baph, b_tph, b_dia, b_ht, b_HD, b_slope, b_aspect, b_sdi) ~ dnorm(0, 4)
  )
  
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod1_fit = quap(mod1, data = boot_data) #, start = list(b0=0, b_qmd=0, b_baph=0, b_tph=0, b_dia=0, b_ht=0, b_HD=0, b_slope=0, b_aspect=0, sigma=1))
        lppd_cv = as.data.frame(sum(lppd_newdata(mod1_fit, newdata = test_data)))
        boolFalse <- T
      }, error=function(e){},finally={}
    )
  }
  names(lppd_cv)="lppd_cv"
  r2_boot=median(bayesian_rsq(mod1_fit, boot_data$sqrtcpa))
  names(r2_boot)="r2_boot"
  r2_out=median(bayesian_rsq_out(mod1_fit, obs = test_data$sqrtcpa, new_values = test_data))
  names(r2_out)="r2_out"
  stat=cbind(lppd_cv, r2_boot, r2_out)
  return(stat)
}
cv_list <- do.call(rbind,cross_v)
median(cv_list$r2_boot)
ggplot() + geom_histogram(aes(cv_list$r2_boot)) + 
		   geom_vline(aes(xintercept=median(cv_list$r2_boot)), linetype="dashed", linewidth=1) +
		   theme_classic()
median(cv_list$r2_out, na.rm = T)
ggplot() + geom_histogram(aes(cv_list$r2_out)) + 
		   geom_vline(aes(xintercept=median(cv_list$r2_out)), linetype="dashed", linewidth=1) +
		   theme_classic()
sum(cv_list$lppd_cv)

```

```{r}
precis(mod1_fit, prob = 0.95)
r_in <- bayesian_rsq(mod1_fit, boot_data$sqrtcpa)
r_out <- bayesian_rsq_out(mod1_fit, obs = test_data$sqrtcpa, new_values = test_data)

ggplot() + geom_histogram(aes(x=r_in)) + 
		   geom_vline(aes(xintercept=median(r_in)), linetype="dashed", linewidth=1) +
		   theme_classic()

ggplot() + geom_histogram(aes(x=r_out)) + 
		   geom_vline(aes(xintercept=median(r_out)), linetype="dashed", linewidth=1) +
		   theme_classic()
```


```{r}
precis(mod1_fit, depth=2)
#psis = PSIS(mod1_fit, depth = 2)
mod1_psis = PSIS(mod1_fit, depth = 2, pointwise = T)
psis

waic = WAIC(mod1_fit, depth = 2)
waic

pred = link(mod1_fit, n = 10000)
mean_pred1 = colMeans(pred)
resid = final_plot_data_subset$cpa - mean_pred1
all_resid = t(final_plot_data_subset$cpa - t(pred))
resid_var = apply(all_resid, 1, var)
mean_var =  apply(pred, 1, var)
R2 = mean_var / (mean_var + resid_var)

pred = link(mod1_fit, n = 10000)
mean_pred1 = colMeans(pred)
resid = test_data$cpa - mean_pred1
all_resid = t(test_data$cpa - t(pred))
resid_var = apply(all_resid, 1, var)
mean_var =  apply(pred, 1, var)
R2 = mean_var / (mean_var + resid_var)

pred = link(mod1_fit, n=10000)
all_resid = t(test_data$cpa - t(pred))
resid_var = apply(all_resid, 1, var)
mean_var =  apply(pred, 1, var)
R2 = mean_var / (mean_var + resid_var)



median(bayesian_rsq(mod1_fit, boot_data))

median(R2)

ggplot() + geom_histogram(aes(x=R2)) + 
		   geom_vline(aes(xintercept=median(R2)), linetype="dashed", linewidth=1) +
		   theme_classic()

ggsave("bayesian_r2.pdf", width=4, height=3)
```

## 1.2 Psuedo-cross validation but with 90/10 split and p-number iterations - psuedo bootstap
```{r}
i = 9

final_plot_data_subset <- final_plot_data %>% 
  filter(ownership == ownership_index[i])# %>% 
  #filter(fortypcd < 600 & fortypcd >= 500)

plot_IDs <- unique(final_plot_data_subset$uniq_spl_unit_id)

sample <- final_plot_data_subset
sample$logcpa <- log(sample$cpa)
sample$sqrtcpa <- (sample$cpa)^0.5

#subset = final_plot_data_subset
#subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] = scale(subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")])

cross_v <- foreach(p = 1:1000, .packages = c("rethinking", "dplyr"))%dopar%{ #    .errorhandling = 'remove'
  
  smp_size <- floor(0.9*length(plot_IDs)) # switch to 0.8 for 80/20 split
  plot_ind <- as.data.frame(sample(plot_IDs, size = smp_size))
  colnames(plot_ind) <- "uniq_spl_unit_id"
  
  boot_data <- inner_join(sample, plot_ind)
  boot_data_raw <- boot_data
  test_data <- anti_join(sample, boot_data)
  
  boot_data = boot_data[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi", "fortypcd")]
  test_data = test_data[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi", "fortypcd")]

  # Specify model
  mod1 = alist(
    cpa ~ dnorm(mu, sigma),
    mu <- b0  + b_qmd*qmd + b_baph*baph + b_tph*tph + b_dia*mean_dia + a[fortypcd] + b_ht*mean_actualht + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    b0 ~ dnorm(100,20),
    c(sigma, sigma_r) ~ dexp(1),
    a[fortypcd] ~ dnorm(0, sigma_r),
    c(b_qmd, b_baph, b_tph, b_dia, b_ht, b_HD, b_slope, b_aspect, b_sdi) ~ dnorm(0, 5)
  )
  
  boot_data$fortypcd = coerce_index(as.factor(boot_data$fortypcd))
  test_data$fortypcd = coerce_index(as.factor(test_data$fortypcd))
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod1_fit = quap(mod1, data = boot_data) #, start = list(b0=0, b_qmd=0, b_baph=0, b_tph=0, b_dia=0, b_ht=0, b_HD=0, b_slope=0, b_aspect=0, sigma=1))
        lppd_cv = as.data.frame(sum(lppd_newdata(mod1_fit, newdata = test_data)))
        names(lppd_cv)="lppd_cv"
        r2_boot=median(bayesian_rsq(mod1_fit, boot_data$cpa))
        names(r2_boot)="r2_boot"
        r2_out=median(bayesian_rsq_out(mod1_fit, obs = test_data$cpa, new_values = test_data))
        names(r2_out)="r2_out"
        stat=cbind(lppd_cv, r2_boot, r2_out)
        boolFalse <- T
      }, error=function(e){},finally={}
    )
  }
  

  return(stat)
}

cv_list <- do.call(rbind,cross_v)
median(cv_list$r2_boot)
ggplot() + geom_histogram(aes(cv_list$r2_boot)) + 
		   geom_vline(aes(xintercept=median(cv_list$r2_boot)), linetype="dashed", linewidth=1) +
		   theme_classic()
median(cv_list$r2_out, na.rm = T)
ggplot() + geom_histogram(aes(cv_list$r2_out)) + 
		   geom_vline(aes(xintercept=median(cv_list$r2_out)), linetype="dashed", linewidth=1) +
		   theme_classic()
sum(cv_list$lppd_cv, na.rm = T)
```

```{r}
precis(mod1_fit, prob = 0.95)
r_in <- bayesian_rsq(mod1_fit, boot_data$sqrtcpa)
r_out <- bayesian_rsq_out(mod1_fit, obs = test_data$sqrtcpa, new_values = test_data)

ggplot() + geom_histogram(aes(x=r_in)) + 
		   geom_vline(aes(xintercept=median(r_in)), linetype="dashed", linewidth=1) +
		   theme_classic()

ggplot() + geom_histogram(aes(x=r_out)) + 
		   geom_vline(aes(xintercept=median(r_out)), linetype="dashed", linewidth=1) +
		   theme_classic()
```

## XX Laplace prior - not usable
```{r}
i = 2

final_plot_data_subset <- final_plot_data %>% 
  filter(ownership == ownership_index[i])

plot_IDs <- unique(final_plot_data_subset$uniq_spl_unit_id)

cross_v <- foreach(p = 1:length(plot_IDs), .packages = c("rethinking"))%dopar%{ #    .errorhandling = 'remove'
  id = plot_IDs[p]
  subset = final_plot_data_subset
  subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] = scale(subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")])
  
  boot_data = subset[which(subset$uniq_spl_unit_id != id),]
  test_data = subset[which(subset$uniq_spl_unit_id == id),]
  
  boot_data = boot_data[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] #"fortypcd",
  test_data = test_data[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")]

  # Specify model
  mod1 = alist(
    cpa ~ dlaplace(mu, sigma),
    mu <- b0  + b_qmd*qmd + b_baph*baph + b_tph*tph + b_dia*mean_dia + #+ alpha[fortypcd]
      b_ht*mean_actualht + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    b0 ~ dlaplace(0, 1),
    sigma ~ dexp(1),
    c(b_qmd, b_baph, b_tph, b_dia, b_dia, b_ht, b_HD, b_slope, b_aspect, b_sdi) ~ dlaplace(0, 2)
  )
  
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod1_fit = quap(mod1, data = boot_data, start = list(b0=0, b_qmd=0, b_baph=0, b_tph=0, b_dia=0, b_ht=0, b_HD=0, b_slope=0, b_aspect=0, sigma=1))
        lppd_cv = as.data.frame(sum(lppd_newdata(mod1_fit, newdata = test_data)))
        boolFalse <- T
      }, error=function(e){},finally={}
    )
  }
  
  names(lppd_cv)="lppd_cv"
  r2_boot=median(bayesian_rsq(mod1_fit, boot_data$cpa))
  names(r2_boot)="r2_boot"
  r2_out=median(bayesian_rsq(mod1_fit, test_data$cpa))
  names(r2_out)="r2_out"
  stat=cbind(lppd_cv, r2_boot, r2_out)
  return(stat)
}
cv_list <- do.call(rbind,cross_v)
mod1_lppd_cv = sum(cv_list$lppd_cv)
```
## 1.2 By ownership, w random effect - no usable 

```{r}
i = 9

stat1 <- data.frame()

final_plot_data_subset <- final_plot_data %>% 
  filter(ownership == ownership_index[i])

plot_IDs <- unique(final_plot_data_subset$uniq_spl_unit_id)

cross_v <- foreach(p = 1:length(plot_IDs), .packages = c("rethinking"))%dopar%{ #    .errorhandling = 'remove'
  id = plot_IDs[p]
  subset = final_plot_data_subset
  # subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] = scale(subset[, c( "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")])
  
  boot_data = subset[which(subset$uniq_spl_unit_id != id),]
  test_data = subset[which(subset$uniq_spl_unit_id == id),]
  
  
  boot_data = boot_data[, c("fortypcd", "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")] #"fortypcd",
  test_data = test_data[, c("fortypcd", "cpa", "qmd", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")]

  mod2 = alist(
    cpa ~ dnorm(mu, sigma),
    mu <- b0 + alpha[fortypcd] + b_qmd*qmd + b_baph*baph + b_tph*tph + b_dia*mean_dia + b_ht*mean_actualht + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    sigma ~ dexp(1),
    b0 ~ dnorm(0, 10),
    alpha[location_id] ~ dnorm(0, sigma_loc),
    c(b_qmd, b_baph, b_tph, b_dia, b_dia, b_ht, b_HD, b_slope, b_aspect, b_sdi) ~ dnorm(0, 10), #
    sigma_loc ~ dexp(1)
  )
  
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod_sample_fit <- quap(mod2, data = boot_data)
        lppd_cv = sum(lppd_newdata(mod_sample_fit, newdata = test_data))
        in_samp <- bayesian_rsq(mod_sample_fit, boot_data$sqrtcpa)
        r_sq <- median(in_samp)
        r_sq_low <- quantile(in_samp, probs = 0.025)
        r_sq_high <- quantile(in_samp, probs = 0.975)
        
        out_samp <- bayesian_rsq_out(mod_sample_fit, test_data$sqrtcpa, test_data)
        r_sq_out <- median(out_samp)
        r_sq_out_low <- quantile(out_samp, probs = 0.025)
        r_sq_out_high <- quantile(out_samp, probs = 0.975)
        boolFalse <- T
      }, error=function(e){}, warning = function(w){}, finally={}
    )
  }
  
  region <- ownership_index[i]
  
  temp <- cbind(region, lppd_cv, r_sq, r_sq_low, r_sq_high, r_sq_out, r_sq_out_low, r_sq_out_high)
  
  stat1 <- rbind(stat1, temp)
}

```




## 2 New approach
### 2.1 Complete R-sq with logcpa and quap and no random effect
```{r}
stat_log = data.frame()

for(i in 1:length(ownership_index)){
  sample <- final_plot_data %>% 
    filter(ownership == ownership_index[i]) %>% 
    filter(fortypcd < 600 & fortypcd >= 500)
  sample$fortypcd <- as.factor(sample$fortypcd)
  sample$logcpa <- log(sample$cpa)
  sample$sqrtcpa <- (sample$cpa)^0.5
  
  plotIDs <- unique(sample$uniq_spl_unit_id)
  smp_size <- floor(0.9*length(plotIDs)) # switch to 0.8 for 80/20 split
  plot_ind <- as.data.frame(sample(plotIDs, size = smp_size))
  colnames(plot_ind) <- "uniq_spl_unit_id"
  
  boot_data <- inner_join(sample, plot_ind)
  test_data <- anti_join(sample, boot_data)
  
  mod_sample = alist(
    logcpa ~ dnorm(mu, sigma),
    mu <- b0 + b_d*mean_dia + b_ht*mean_actualht + b_tph*tph + b_ba*baph + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    sigma ~ dexp(.5),
    b0 ~ dnorm(5, 4),  
    c(b_d, b_ht, b_tph, b_ba, b_HD, b_slope, b_aspect, b_sdi) ~ dnorm(0, 4)
  )
  
  stan_data = boot_data[, c("logcpa", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")]
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod_sample_fit <- quap(mod_sample, data = stan_data)
        lppd_cv = sum(lppd_newdata(mod1_fit, newdata = test_data))
        
        in_samp <- bayesian_rsq(mod_sample_fit, boot_data$logcpa)
        r_sq <- median(in_samp)
        r_sq_low <- quantile(in_samp, probs = 0.025)
        r_sq_high <- quantile(in_samp, probs = 0.975)
        
        out_samp <- bayesian_rsq_out(mod_sample_fit, test_data$logcpa, test_data)
        r_sq_out <- median(out_samp)
        r_sq_out_low <- quantile(out_samp, probs = 0.025)
        r_sq_out_high <- quantile(out_samp, probs = 0.975)
        boolFalse <- T
      }, error=function(e){}, warning = function(w){}, finally={}
    )
  }
  
  region <- ownership_index[i]
  
  temp <- cbind(region, lppd_cv, r_sq, r_sq_low, r_sq_high, r_sq_out, r_sq_out_low, r_sq_out_high)
  
  stat_log <- rbind(stat_log, temp)
}

```


### 2.2 Complete R-sq with sqrtcpa with quap no random effect
```{r}
stat_sqrt = data.frame()

for(i in 1:length(ownership_index)){
  sample <- final_plot_data %>% 
    filter(ownership == ownership_index[i]) %>% 
    filter(fortypcd < 600 & fortypcd >= 500)
  sample$fortypcd <- as.factor(sample$fortypcd)
  sample$logcpa <- log(sample$cpa)
  sample$sqrtcpa <- (sample$cpa)^0.5
  
  plotIDs <- unique(sample$uniq_spl_unit_id)
  smp_size <- floor(0.90*length(plotIDs)) # switch to 0.8 for 80/20 split
  plot_ind <- as.data.frame(sample(plotIDs, size = smp_size))
  colnames(plot_ind) <- "uniq_spl_unit_id"
  
  boot_data <- inner_join(sample, plot_ind)
  test_data <- anti_join(sample, boot_data)
  
  mod_sample = alist(
    sqrtcpa ~ dnorm(mu, sigma),
    mu <- b0 + b_d*mean_dia + b_ht*mean_actualht + b_tph*tph + b_ba*baph + b_HD*mean_HD + b_slope*slope + b_aspect*aspect + b_sdi*sdi,
    sigma ~ dexp(.5),
    b0 ~ dnorm(8, 5),  
    c(b_d, b_ht, b_tph, b_ba, b_HD, b_slope, b_aspect, b_sdi) ~ dnorm(0, 4)
  )
  
  stan_data = boot_data[, c("sqrtcpa", "baph", "tph", "mean_dia", "mean_actualht", "mean_HD", "slope", "aspect", "sdi")]
  
  boolFalse <- F
  while(boolFalse == F){
    tryCatch(
      {
        mod_sample_fit <- quap(mod_sample, data = stan_data)
        lppd_cv = sum(lppd_newdata(mod1_fit, newdata = test_data))
        
        in_samp <- bayesian_rsq(mod_sample_fit, boot_data$sqrtcpa)
        r_sq <- median(in_samp)
        r_sq_low <- quantile(in_samp, probs = 0.025)
        r_sq_high <- quantile(in_samp, probs = 0.975)
        
        out_samp <- bayesian_rsq_out(mod_sample_fit, test_data$sqrtcpa, test_data)
        r_sq_out <- median(out_samp)
        r_sq_out_low <- quantile(out_samp, probs = 0.025)
        r_sq_out_high <- quantile(out_samp, probs = 0.975)
        boolFalse <- T
      }, error=function(e){}, warning = function(w){}, finally={}
    )
  }
  
  region <- ownership_index[i]
  
  temp <- cbind(region, lppd_cv, r_sq, r_sq_low, r_sq_high, r_sq_out, r_sq_out_low, r_sq_out_high)
  
  stat_sqrt <- rbind(stat_sqrt, temp)
}
```

## 3.0 MCMC fit

```{r}
i=9

sample <- final_plot_data %>% 
  filter(ownership == ownership_index[i]) %>% 
  filter(fortypcd < 600 & fortypcd >= 500)
forest_type = unique(sample$fortypcd)
sample$fortypcd <- coerce_index(as.factor(sample$fortypcd))
sample$logcpa <- log(sample$cpa)
sample$sqrtcpa <- (sample$cpa)^0.5

plotIDs <- unique(sample$uniq_spl_unit_id)
smp_size <- floor(0.90*length(plotIDs)) # switch to 0.8 for 80/20 split
plot_ind <- as.data.frame(sample(plotIDs, size = smp_size))
colnames(plot_ind) <- "uniq_spl_unit_id"

boot_data <- inner_join(sample, plot_ind)
boot_data_raw <- boot_data
test_data <- anti_join(sample, boot_data)

boot_data = boot_data[, c("sqrtcpa", "mean_dia", "mean_actualht", "fortypcd")]
test_data = test_data[, c("sqrtcpa", "mean_dia", "mean_actualht", "fortypcd")]
  
mod2 = alist(
    sqrtcpa ~ dnorm(mu, sigma),
    mu <- b0 + b_d*mean_dia + b_ht*mean_actualht + alpha[fortypcd],
    sigma ~ dexp(1),
    b0 ~ dnorm(10, 4),
    alpha[fortypcd] ~ dnorm(0, sigma_loc),
    c(b_d, b_ht) ~ dnorm(0, 4),
    sigma_loc ~ dexp(1)
  )

pmod_fit1 = ulam(mod2, data = boot_data, iter = 2000, warmup = 500, chains = 4, cores = 6, log_lik = T)

traceplot(pmod_fit1)
trankplot(pmod_fit1)
precis(pmod_fit1)
precis(pmod_fit1, depth=2, pars=c("alpha"))

post = extract.samples(pmod_fit1)
alpha_post = post$alpha + as.vector(post$b0) + as.vector(post$b_d) + as.vector(post$b_ht)
colnames(alpha_post) = paste0("forest_type_", forest_type)
alpha_dt = reshape2::melt(alpha_post)
colnames(alpha_dt) = c("id", "fortypcd", "value")
alpha_dt$valuesq = (alpha_dt$value)
hist(alpha_dt$valuesq)

ggplot(alpha_dt) + geom_density(aes(x=valuesq, fill=fortypcd)) + theme_classic() +
				   xlab("Predicted sqrt(C) (sqrt(MgC/HA))")
```

```{r}
hist(boot_data_raw$cpa) 
hist(sqrt(boot_data_raw$cpa))
hist(sample$cpa) 
hist(sqrt(sample$cpa)) 
hist(log(sample$cpa)) 
```

```{r}
post_alpha <- post$alpha
alpha_mean = colMeans(post_alpha)
qqnorm(alpha_mean)
qqline(alpha_mean)
```

```{r}
postcheck(pmod_fit1)
```

```{r}
precis(pmod_fit1, prob = 0.95)
WAIC(mod1_fit)
WAIC(pmod_fit1)
compare(mod1_fit, pmod_fit1, func = WAIC)
compare(mod1_fit, pmod_fit1, func = PSIS)
sum(lppd(pmod_fit1))
PSIS(pmod_fit1, pointwise = T)
```

```{r}
mod1_pred = link(pmod_fit1)

mpv =  as.vector(mod1_pred)

post <- extract.samples(pmod_fit1)                # [web:4]
sigma_draws <- as.vector(post$sigma)

# draw normal predictive samples: one y_rep per (sample, case)
mod1_pred_counts <- array(
  rnorm(length(mpv), mean = mpv, sd = rep(sigma_draws, times = ncol(mod1_pred))),
  dim = dim(mod1_pred)
)

# Compare the observed and predicted distribution of data
# y : Takes the observed data
# yrep " Takes in posterior samples of the predicted data
# This function then compares the two distributions
p1 <- bayesplot::ppc_dens_overlay(
  y    = boot_data$sqrtcpa,
  yrep = mod1_pred_counts[1:500, , drop = FALSE])

p1 + theme_classic() + xlab("sqrt cpa") + ylab("Density")
#ggsave("obs_pred_predictions_ranef.pdf", width=4, height=3)
```

Testing 
```{r}
mod1_pred = link(pmod_fit1, data = test_data,  n=10000)
mpv =  as.vector(mod1_pred)

post <- extract.samples(pmod_fit1)                # [web:4]
sigma_draws <- as.vector(post$sigma)

# draw normal predictive samples: one y_rep per (sample, case)
mod1_pred_counts <- array(
  rnorm(length(mpv), mean = mpv, sd = rep(sigma_draws, times = ncol(mod1_pred))),
  dim = dim(mod1_pred)
)

# Compare the observed and predicted distribution of data
# y : Takes the observed data
# yrep " Takes in posterior samples of the predicted data
# This function then compares the two distributions
p1 <- bayesplot::ppc_dens_overlay(
  y    = test_data$sqrtcpa,
  yrep = mod1_pred_counts[1:500, , drop = FALSE])

p1 + theme_classic() + xlab("sqrt cpa") + ylab("Density")
#ggsave("obs_pred_predictions_ranef.pdf", width=4, height=3)
```