---
title: "Buffering data comparison"
author: "Tara Skiba"
date: "2026-03-23"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comparing raw, public, and buffered public coordinate values (10 buffered with r = 5280 ft)

Using WorldClim Bio data (pixel size: 927.67 m or 3043.537 ft)

```{r library}
library(readxl)
library(dplyr)
library(Metrics)
library(maditr)
library(doParallel)
library(tidyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(stats)

no_cores <- detectCores() - 3
cl <- makeCluster(no_cores, type = "PSOCK")
registerDoParallel(cl)
```
 ## Splitting dataset of 1 buffered coordinate to run through *skiba*
```{r}

point.split <- read.csv("5280.0ftsingle.csv")

```

 ## Splitting dataset of 10 buffered coordinates to run through *skiba*
```{r}

point.split <- read.csv("5280.0ft10points.csv")
26940/2 # end of 51_165_93
point.split.1 <- point.split[1:13470,]
point.split.2 <- point.split[13470:26940,]

write.csv(point.split.1, "split1.csv")
write.csv(point.split.2, "split2.csv")
```

```{r datafiles}
raw <- read.csv("WORLDCLIM_V1_BIO_RAW.csv")

public <- read.csv("WORLDCLIM_V1_BIO_PUBLIC.csv")

buffer1 <- read.csv("WORLDCLIM_V1_BIO_SINGLE_BUFFER.csv")
buffer1 <- buffer1[, -c(21,22)]

split1 <- read.csv("WORLDCLIM_V1_BIO_BUFFER1.csv")
split2 <- read.csv("WORLDCLIM_V1_BIO_BUFFER2.csv")
buffer10 <- bind_rows(split1, split2)

final_plot_data <- read.csv("tree_plot_data_final_carbon.csv")
final_plot_data <- final_plot_data %>% 
  filter(ownership != "New River Gorge")
final_plot_data$plot_ID <- paste(final_plot_data$STATECD, 
                                 final_plot_data$COUNTYCD, 
                                 final_plot_data$PLOT,
                                sep = "_")

```

# Difference calcaulations
```{r}
set.seed = 123
variables <- c(colnames(buffer1[,-20]))

difference_calculations <- function(dataset, bio_var){
  
    percent.diff <- 100*(dataset[,1]-dataset[,2])/dataset[,2] # as a percent
    percent.diff[sapply(percent.diff, is.infinite)] <- NA
    percent.diff <- mean(percent.diff, na.rm = T)
        
    abs.diff <- abs(dataset[,2]-dataset[,1])
    abs.diff <- mean(abs.diff, na.rm = T)
    
    df <- data.frame(biovar = bio_var, percent.diff = percent.diff, abs.diff = abs.diff)
    return(df)
    
}

data_cleaning <- function(dataset){
  dataset = dataset[complete.cases(dataset),]
  dataset = dataset[,order(colnames(dataset))]
  return(dataset)
}

differences <- foreach(var = 1:length(variables), .packages = c("dplyr"))%dopar%{
    var.pub = public[,c(variables[var], "plot_ID")]
    var.raw = raw[,c(variables[var], "plot_ID")]
    var.buf1 = buffer1[,c(variables[var], "plot_ID")]
    var.buf10 = buffer10[,c(variables[var], "plot_ID")]
   
    # pub/raw (PR)
    join.pub.raw = merge(var.pub, var.raw, by = "plot_ID")
    join.pub.raw = data_cleaning(join.pub.raw)
    # buf1/raw
    join.buf1.raw = merge(var.buf1, var.raw, by = "plot_ID")
    join.buf1.raw = data_cleaning(join.buf1.raw)
    # buf10/raw
    join.buf10.raw = merge(var.buf10, var.raw, by = "plot_ID")
    join.buf10.raw = data_cleaning(join.buf10.raw)
    #buff10 vs buf1
    join.buf10.buf1 = merge(var.buf10, var.buf1, by = "plot_ID")
    join.buf10.buf1 = data_cleaning(join.buf10.buf1)
    
    
    PR <- difference_calculations(join.pub.raw, var)
    PR$compare = "PR"
    B1R <- difference_calculations(join.buf1.raw, var)
    B1R$compare = "B1R"
    B10R <- difference_calculations(join.buf10.raw, var)
    B10R$compare = "B10R"
    B10B1 <- difference_calculations(join.buf10.buf1, var)
    B10B1$compare = "B10B1"
    
    diff <- rbind(PR, B1R, B10R, B10B1)
    
    return(diff)
}
diff.bind <- do.call(rbind, differences)
diff.bind

```

## Percent difference
```{r}
percent.diff.tab <- diff.bind[,-c(3)]
percent.diff.tab <- percent.diff.tab %>% 
  pivot_wider(names_from = compare, values_from = percent.diff)
clipr::write_clip(percent.diff.tab)
percent.diff.tab
```
## Abs difference
```{r}
abs.diff.tab <- diff.bind[,-c(2)]
abs.diff.tab <- abs.diff.tab %>% 
  pivot_wider(names_from = compare, values_from = abs.diff)
clipr::write_clip(abs.diff.tab)
abs.diff.tab
```

# CV and CI for 10-point buffered coordinates.
```{r}
buff.nonagg.1 <- read.csv("WORLDCLIM_V1_BIO_BUFFNONAGG1.csv")
buff.nonagg.2 <- read.csv("WORLDCLIM_V1_BIO_BUFFNONAGG2.csv")
buff.nonagg <- bind_rows(buff.nonagg.1, buff.nonagg.2)
summary(buff.nonagg)
clipr::write_clip(summary(buff.nonagg))
```

```{r}
buff.stats <- setNames(data.frame(matrix(ncol = 2, nrow = 0)), c("var", "cv"))
# buf.stats <- data.frame(matrix(ncol = 2, nrow = 0))
# x <- c("var", "cv")
# colnames(buf.stats) <- x
for(i in 1:length(variables)){
  data.var <- buff.nonagg %>% 
    select(c(plot_ID, matches(variables[i]))) %>% 
    dplyr::group_by(plot_ID) %>% 
    dplyr::summarise_all(.funs= c("sd"=~sd(.),
                                 "mean"=~mean(.)),
                       na.rm = T)
  data.var$cv = data.var$sd/data.var$mean
  data.var <- data.var[is.finite(data.var$cv),]
  cv <- mean(data.var$cv, na.rm = T)*100
  var <- variables[i]
  cv.var <- cbind(var,cv)
  buff.stats <- rbind(buff.stats, cv.var)
}

buff.stats
clipr::write_clip(buff.stats)
```

# Visuals of 10-point buffered approach

```{r}

ggplot(buff.nonagg, aes(x=bio12)) +
geom_histogram(aes(y=..density..), position="identity", alpha=0.5)+
geom_density(alpha=0.6)+
labs(title="bio12 Histogram",x="Annual Precipitation (mm)", y = "Density")+
theme_classic()

ggplot(buff.nonagg, aes(x=bio15)) +
geom_histogram(aes(y=..density..), position="identity", alpha=0.5)+
geom_density(alpha=0.6)+
labs(title="bio15 Histogram",x="Seasonal Precipitation (cv)", y = "Density")+
theme_classic()
```

# Elevation comparison

```{r}
fia <- read_xlsx("fed forest cond data 04-03-2026 no coor.xlsx") 
fia$plot_ID <- paste(fia$STATECD, 
                     fia$COUNTYCD, 
                     fia$PLOT,
                     sep = "_")
fia <- fia %>% group_by(plot_ID) %>% top_n(1, MEASYEAR)
fia <- unique(fia[,c("plot_ID", "ELEV")])
dep <- read.csv("USGS_3DEP_1m_raw.csv")
dep$dep <- dep$first*3.280839895
dep <- dep[,c("plot_ID", "dep")]
srtmgl <- read.csv("USGS_SRTMGL1_003_raw.csv") 
srtmgl$srtmgl <- srtmgl$first*3.280839895
srtmgl <- srtmgl[,c("plot_ID", "srtmgl")]

elevation <- full_join(dep, srtmgl)
elevation <- left_join(elevation, fia)

dep_diff <- mean(100*(elevation[,"dep"]-elevation[,"ELEV"])/elevation[,"ELEV"], na.rm = T)
srtmgl_diff <- mean(100*(elevation[,"srtmgl"]-elevation[,"ELEV"])/elevation[,"ELEV"], na.rm = T)

abs.diff <- abs(elevation[,"srtmgl"]-elevation[,"ELEV"])
abs.diff <- mean(abs.diff, na.rm = T)
dep_diff
print("srtmgl_diff: ")
srtmgl_diff
print("abs.diff")
abs.diff
```
```{r}
require(gridExtra)
plot1 <- ggplot(elevation, aes(x=ELEV)) + geom_histogram(binwidth=200, color="black", fill="white") + xlab("FIA (ft)") + ylim(0,250) + theme_minimal()
plot2 <- ggplot(elevation, aes(x=srtmgl)) + geom_histogram(binwidth=200,color="black", fill="white") + xlab("NASA (ft)") + ylim(0,250) + theme_minimal()
grid.arrange(plot1, plot2, ncol=2, top = textGrob("Distribution of Elevation", gp = gpar(fontsize = 20, font = 3)))

ggplot(elevation, aes(x = ELEV, y = srtmgl)) + geom_point() + labs(title = "Queried vs Recorded Elevation Values") + xlab("FIA (ft)") + ylab("NASA (ft)") + geom_abline(intercept = 0, slope = 1, colour = "red") + theme_bw()

#some outliers are driving the differences up

elevation %>% summarise_at(.vars = c("ELEV", "srtmgl"),
                           .funs = c(med = ~median(.),
                                      mean = ~mean(.)))

```
##  Data pruning of outliers
```{r}
# as according to Faraway et al. book

elevation_cleaned <- elevation
elevation_cleaned$abs_diff = abs(elevation_cleaned$srtmgl-elevation_cleaned$ELEV)
summary(elevation_cleaned$abs_diff)
precleaned_elevation_plot = plot(elevation_cleaned$abs_diff)
length(elevation_cleaned$abs_diff)

#remove outliers greater than 3000 - too conservative - only removed 2
#remove outliers greater than 2000 - too conservative
elevation_cleaned <- elevation_cleaned %>% 
  filter(abs_diff < 2000)
cleaned_elevation_plot = plot(elevation_cleaned$abs_diff)


length(elevation_cleaned)

# continuing to filter the data set to include abs diff less than 300 (the 3rd quartile), the values were only reduced to % diff of 1.652476 and abs diff of 119.9826 ft
```

## refit statistics
```{r}

srtmgl_diff2 <- mean(100*(elevation_cleaned[,"srtmgl"]-elevation_cleaned[,"ELEV"])/elevation_cleaned[,"ELEV"], na.rm = T)

abs.diff2 <- abs(elevation_cleaned[,"srtmgl"]-elevation_cleaned[,"ELEV"])
abs.diff2 <- mean(abs.diff2, na.rm = T)

srtmgl_diff2
abs.diff2

```

## fit as model for fun
```{r}
model1 <- lm(ELEV ~ srtmgl, data = elevation)
summary(model1)
plot(model1)

model2 <- lm(ELEV ~ srtmgl, data = elevation_cleaned)
summary(model2)
plot(model2)

cor(elevation$srtmgl, elevation$ELEV)
cor(elevation_cleaned$srtmgl, elevation_cleaned$ELEV)
```
keep all points in. 1-to-1 plot and correlation coefficient. don't include data cleaning
